\LARGE \bf Expectation-Maximization based defense mechanism for distributed Model Predictive Control

### 2022-05-09 article/article.tex
function to colon operation

\emph{Notation:}
In this paper, $\norm{\cdot}$, $\norm{\cdot}_{F}$, and ${\norm{\vec{v}}_{Y}=\norm{Y^{\frac{1}{2}}\vec{v}}}$ represent the $\ell_{2}$, Frobenius and weighted norms.
$\Proj^{\set{T}}(\cdot)$ is the Euclidean projection onto $\set{T}$.
% $\card{x}$ is the number of elements in $x$.
$\kron{}{}$ ($\hadamard{}{}$) is the Kronecker (Hadamard) product.
$\vectorize{A}$ vectorizes matrix $A$.
${n\mathbin{:}i\mathbin{:}j}$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where ${m=\mathrm{truncate}(\frac{j-n}{i})}$, and ${n\mathbin{:}j}$ means ${i=1}$.
% $[A,B]$ ($[A;B]$) is horizontal (vertical) concatenation of matrices.
${\0_{m}=0_{m\times 1}}$ (${\1_{m}=1_{m\times 1}}$), where
$0_{m\times n}$ ($1_{m\times n}$) is a $0$ ($1$) filled \mbox{$m$-by-$n$} matrix.
$\indicator{x}$ is the indicator function returning $1$ if $x$ is true and $0$ otherwise.
$\symmetric^{n}$, and $\defpos^{n}$ ($\semidefpos^{n}$) are symmetric and positive (semi-)definite matrices of size $n$.
${A^{\dagger}={{(A^{T}A)}^{-1}A^{T}}}$.
% $\determinant{A}$ is the determinant of $A$.
A vector $\vec{v}_{i}$, correspond to the $i$-th agent, and these vectors can be stacked in a vector $\vec{v}$.
% $\random{a},\random{A},\randomvec{a}$ are random variables, scalar, matrix and vector respectively.
$\expectation{\random{x}}$ is the expected value of random variable $\random{x}$, and $\probability{A \mid B}$ is the conditional probability of $A$ given condition $B$.
% $\pdf{\cdot}$ is a probability density function.
${\diag(A_{1},\dots,A_{N})}$ corresponds to a block diagonal matrix.

@@

\subsection{Model Predictive Control}\label{ssec:MPC}
Consider a system composed of discrete-time linear time-invariant agents ${i\in\set{M}=\{1\mathbin{:}\nsubsystems\}}$, modeled by
\begin{equation}
\begin{matrix}
  \label{eq:systems}
\vec{x}_{i}[k+1]&=&A_{i}\xik &+& B_{i}\uik\\
\vec{y}_{i}[k+1]&=&C_{i}\xik &&
\end{matrix},
\end{equation}
with ${\vec{x}_{i}[k]\in R^{n_{x}}}$, ${\vec{u}_{i}[k]\in R^{n_{u}}}$ and ${\vec{y}_{i}[k]\in R^{n_{y}}}$.

@@

The overall system is controlled by a \mpc{}, which
computes the optimal input for a finite prediction horizon ${\set{H}=\{1\mathbin{:}\predhorz\}}$ by solving the following problem:
\begin{equation}\label{eq:GOP}
\resizebox{.88\columnwidth}{!}{$
\begin{matrix}
\underset{\useq}{\mathrm{minimize}}&\resizebox{0.35\textwidth}{!}{$\overbrace{\sum\limits_{i\in\set{M}} \overbrace{\sum_{j\in\set{H}}\norm{\mpcvec{v}[i][k+j][k]}^{2}_{Q_i}+\norm{\mpcvec{u}[i][k+j-1][k]}^{2}_{R_i}}^{\textstyle{} \obji}}^{\textstyle{} \globobj}$}\\
\mathrm{subject~ to}&~\eqref{eq:systems},\eqref{eq:u_local_constraints}\ \mathrm{and}\ \eqref{eq:global_constraint}
\left\}\small
\begin{aligned}
  &\forall i\in\set{M}\\
  &\forall j\in\set{H}
\end{aligned}\right.,

@@

We can get an explicit solution for the dual variables $\vec{\lambda}_{i}[k]$, which are \acrfull{pwa} functions w.r.t. $\vec{\theta}_{i}[k]$:
% \begin{equation}
%   \resizebox{.88\columnwidth}{!}{$
%     \begin{aligned}
%       \label{eq:lambdafuntheta}
%       \lambdaik=
%       \begin{cases}
%         -P_{i}^{1}\thetaik-\vec{s}_{i}^{1}[k]&\text{if}\ G^{1}[k]\thetaik \preceq b^{1}[k] \\
%         \qquad\quad \vdots&\qquad\quad \vdots\\
%         -P_{i}^{N}\thetaik-\vec{s}_{i}^{N}[k]&\text{if}\ G^{N}[k]\thetaik \preceq b^{N}[k] \\
%       \end{cases}
%     \end{aligned}
%     $}
% \end{equation}
\begin{equation}
\resizebox{.88\columnwidth}{!}{$
  \begin{aligned}\label{eq:lambdafuntheta}
    \lambdaik=
    % \begin{cases}
      -P_{i}^{n}\thetaik-\vec{s}_{i}^{n}[k]\text{, if}\ G^{n}[k]\thetaik \preceq b^{n}[k],
    % \end{cases}
  \end{aligned}
  $}
\end{equation}
with $n\in\{1\mathbin{:}N\}$.

@@

Such attack function results in a \pwa{} solution for $\lambdaicheat[k]$:
% \begin{equation}
%   \resizebox{.88\columnwidth}{!}{$
%     \label{eq:linear_cheating}
%     \lambdaicheat[k]=
%     \begin{cases}
%       -\Tik P_{i}^{1}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{1}[k]&\text{if}\ G_{1}[k]\thetai \preceq b_{1}[k] \\
%       \qquad\quad \vdots&\qquad\quad \vdots\\
%       -\Tik P_{i}^{N}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetai \preceq b_{N}[k],
%     \end{cases}
%     $}
% \end{equation}
\begin{equation}
  \resizebox{.88\columnwidth}{!}{$
    \label{eq:linear_cheating}
    \lambdaicheat[k]=
    % \begin{cases}
      -\tilde{P}_{i}^{n}[k]\thetaik-\tilde{\vec{s}}_{i}^{n}[k]\text{, if}\ G^{n}[k]\thetaik \preceq b^{n}[k],
    % \end{cases}
    $}
\end{equation}
with $n\in\{1\mathbin{:}N\}$, $\tilde{P}_{i}^{n}[k]=\Tik P_{i}^{n}$ and $\tilde{\vec{s}}_{i}^{n}[k]=\Tik\vec{s}_{i}^{n}[k]$.
Observe that as the function is applied only in $\lambdaik$, it does not affect the zones' hyperplanes.

@@

For a group of $O$ exchanges between an agent and the coordinator, we observe the input and response variables, identified as  ${\random{\vec{\theta}}_{o}}$ and ${\random{\vec{\lambda}}_{o}}$, ${o\in\set{O}=\{1\mathbin{:}O\}}$, which can be organized in ${\random{\Theta},\random{\Lambda}\in\R^{c\times O}}$. ${(\random{\Theta},\random{\Lambda})}$ is our set of observable data $\set{B}$.

As~\eqref{eq:linear_cheating} gives us a multidimensional \pwa{} function, we propose using a multidimensional expansion of the model referred to as \emph{mixture of linear regressions} in~\cite{FariaSoromenho2010} to map the relation between $\random{\Theta}$ and $\random{\Lambda}$. We will call the model \emph{mixture of affine regressions}, since our regressors have linear and constant terms:
% TODO(accacio): change subscript n
% \begin{equation}\label{eq:linear_cheating_random}
%   \randomvec{\lambda}_{o}=
%   \begin{cases}
%     -\tilde{P}^{1}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{1}&\text{if in zone } 1\\
%     \qquad\quad \vdots&\qquad\quad \vdots\\
%     -\tilde{P}^{Z}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{Z}&\text{if in zone } Z\\
%   \end{cases}.
% \end{equation}
\begin{equation}\label{eq:linear_cheating_random}
  \randomvec{\lambda}_{o}=
    -\tilde{P}^{z}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{z}\text{, if in zone } z,\\
\end{equation}
with  ${z\in\set{Z}=\{1\mathbin{:}Z\}}$.

>>>>

\emph{Notation:}
In this paper, $\norm{\cdot}$, $\norm{\cdot}_{F}$, and ${\norm{\vec{v}}_{Y}=\norm{Y^{\frac{1}{2}}\vec{v}}}$ represent the $\ell_{2}$, Frobenius and weighted norms.
$\Proj^{\set{T}}(\cdot)$ is the Euclidean projection onto $\set{T}$.
% $\card{x}$ is the number of elements in $x$.
$\kron{}{}$ ($\hadamard{}{}$) is the Kronecker (Hadamard) product.
$\vectorize{A}$ vectorizes matrix $A$.
${n\until i\until j}$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where ${m=\mathrm{truncate}(\frac{j-n}{i})}$, and ${n\until j}$ means ${i=1}$.
% $[A,B]$ ($[A;B]$) is horizontal (vertical) concatenation of matrices.
${\0_{m}=0_{m\times 1}}$ (${\1_{m}=1_{m\times 1}}$), where
$0_{m\times n}$ ($1_{m\times n}$) is a $0$ ($1$) filled \mbox{$m$-by-$n$} matrix.
$\indicator{x}$ is the indicator function returning $1$ if $x$ is true and $0$ otherwise.
$\symmetric^{n}$, and $\defpos^{n}$ ($\semidefpos^{n}$) are symmetric and positive (semi-)definite matrices of size $n$.
${A^{\dagger}={{(A^{T}A)}^{-1}A^{T}}}$.
% $\determinant{A}$ is the determinant of $A$.
A vector $\vec{v}_{i}$, correspond to the $i$-th agent, and these vectors can be stacked in a vector $\vec{v}$.
% $\random{a},\random{A},\randomvec{a}$ are random variables, scalar, matrix and vector respectively.
$\expectation{\random{x}}$ is the expected value of random variable $\random{x}$, and $\probability{A \mid B}$ is the conditional probability of $A$ given condition $B$.
% $\pdf{\cdot}$ is a probability density function.
${\diag(A_{1},\dots,A_{N})}$ corresponds to a block diagonal matrix.

@@

\subsection{Model Predictive Control}\label{ssec:MPC}
Consider a system composed of discrete-time linear time-invariant agents ${i\in\set{M}=\{1\until \nsubsystems\}}$, modeled by
\begin{equation}
\begin{matrix}
  \label{eq:systems}
\vec{x}_{i}[k+1]&=&A_{i}\xik &+& B_{i}\uik\\
\vec{y}_{i}[k+1]&=&C_{i}\xik &&
\end{matrix},
\end{equation}
with ${\vec{x}_{i}[k]\in R^{n_{x}}}$, ${\vec{u}_{i}[k]\in R^{n_{u}}}$ and ${\vec{y}_{i}[k]\in R^{n_{y}}}$.

@@

The overall system is controlled by a \mpc{}, which
computes the optimal input for a finite prediction horizon ${\set{H}=\{1\until \predhorz\}}$ by solving the following problem:
\begin{equation}\label{eq:GOP}
\resizebox{.88\columnwidth}{!}{$
\begin{matrix}
\underset{\useq}{\mathrm{minimize}}&\resizebox{0.35\textwidth}{!}{$\overbrace{\sum\limits_{i\in\set{M}} \overbrace{\sum_{j\in\set{H}}\norm{\mpcvec{v}[i][k+j][k]}^{2}_{Q_i}+\norm{\mpcvec{u}[i][k+j-1][k]}^{2}_{R_i}}^{\textstyle{} \obji}}^{\textstyle{} \globobj}$}\\
\mathrm{subject~ to}&~\eqref{eq:systems},\eqref{eq:u_local_constraints}\ \mathrm{and}\ \eqref{eq:global_constraint}
\left\}\small
\begin{aligned}
  &\forall i\in\set{M}\\
  &\forall j\in\set{H}
\end{aligned}\right.,

@@

We can get an explicit solution for the dual variables $\vec{\lambda}_{i}[k]$, which are \acrfull{pwa} functions w.r.t. $\vec{\theta}_{i}[k]$:
% \begin{equation}
%   \resizebox{.88\columnwidth}{!}{$
%     \begin{aligned}
%       \label{eq:lambdafuntheta}
%       \lambdaik=
%       \begin{cases}
%         -P_{i}^{1}\thetaik-\vec{s}_{i}^{1}[k]&\text{if}\ G^{1}[k]\thetaik \preceq b^{1}[k] \\
%         \qquad\quad \vdots&\qquad\quad \vdots\\
%         -P_{i}^{N}\thetaik-\vec{s}_{i}^{N}[k]&\text{if}\ G^{N}[k]\thetaik \preceq b^{N}[k] \\
%       \end{cases}
%     \end{aligned}
%     $}
% \end{equation}
\begin{equation}
\resizebox{.88\columnwidth}{!}{$
  \begin{aligned}\label{eq:lambdafuntheta}
    \lambdaik=
    % \begin{cases}
      -P_{i}^{n}\thetaik-\vec{s}_{i}^{n}[k]\text{, if}\ G^{n}[k]\thetaik \preceq b^{n}[k],
    % \end{cases}
  \end{aligned}
  $}
\end{equation}
with $n\in\{1\until N\}$.

@@

Such attack function results in a \pwa{} solution for $\lambdaicheat[k]$:
% \begin{equation}
%   \resizebox{.88\columnwidth}{!}{$
%     \label{eq:linear_cheating}
%     \lambdaicheat[k]=
%     \begin{cases}
%       -\Tik P_{i}^{1}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{1}[k]&\text{if}\ G_{1}[k]\thetai \preceq b_{1}[k] \\
%       \qquad\quad \vdots&\qquad\quad \vdots\\
%       -\Tik P_{i}^{N}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetai \preceq b_{N}[k],
%     \end{cases}
%     $}
% \end{equation}
\begin{equation}
  \resizebox{.88\columnwidth}{!}{$
    \label{eq:linear_cheating}
    \lambdaicheat[k]=
    % \begin{cases}
      -\tilde{P}_{i}^{n}[k]\thetaik-\tilde{\vec{s}}_{i}^{n}[k]\text{, if}\ G^{n}[k]\thetaik \preceq b^{n}[k],
    % \end{cases}
    $}
\end{equation}
with $n\in\{1\until N\}$, $\tilde{P}_{i}^{n}[k]=\Tik P_{i}^{n}$ and $\tilde{\vec{s}}_{i}^{n}[k]=\Tik\vec{s}_{i}^{n}[k]$.
Observe that as the function is applied only in $\lambdaik$, it does not affect the zones' hyperplanes.

@@

For a group of $O$ exchanges between an agent and the coordinator, we observe the input and response variables, identified as  ${\random{\vec{\theta}}_{o}}$ and ${\random{\vec{\lambda}}_{o}}$, ${o\in\set{O}=\{1\until O\}}$, which can be organized in ${\random{\Theta},\random{\Lambda}\in\R^{c\times O}}$. ${(\random{\Theta},\random{\Lambda})}$ is our set of observable data $\set{B}$.

As~\eqref{eq:linear_cheating} gives us a multidimensional \pwa{} function, we propose using a multidimensional expansion of the model referred to as \emph{mixture of linear regressions} in~\cite{FariaSoromenho2010} to map the relation between $\random{\Theta}$ and $\random{\Lambda}$. We will call the model \emph{mixture of affine regressions}, since our regressors have linear and constant terms:
% TODO(accacio): change subscript n
% \begin{equation}\label{eq:linear_cheating_random}
%   \randomvec{\lambda}_{o}=
%   \begin{cases}
%     -\tilde{P}^{1}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{1}&\text{if in zone } 1\\
%     \qquad\quad \vdots&\qquad\quad \vdots\\
%     -\tilde{P}^{Z}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{Z}&\text{if in zone } Z\\
%   \end{cases}.
% \end{equation}
\begin{equation}\label{eq:linear_cheating_random}
  \randomvec{\lambda}_{o}=
    -\tilde{P}^{z}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{z}\text{, if in zone } z,\\
\end{equation}
with  ${z\in\set{Z}=\{1\until Z\}}$.
